---
title: "Labor Data Report for Will Jones"
author: "Jake Carlson"
header-includes:
   - \usepackage{longtable}
output: pdf_document
---
```{r, echo=FALSE}
targetType <- "latex"
```
```{r, echo=FALSE, results="hide", message=FALSE, warning=FALSE, fig.show="hide"}  
source("UnionAnalysis.R")  
```

This report contains graphs and tables from analysis of the Current Population Survey from 1962-2014, using March surveys. 

##Median Incomes by Demographic Categories
To attempt to replicate as closely as possible the estimates from the Census Bureau, median incomes are for those 15 years old and older, and those who have greater than zero dollars in total income. This does not perfectly replicate the Census estimates, but comes close. Incomes have been adjusted to 2008 dollars.  
    
```{r, echo=FALSE, results='asis'}
tabMedIncWK <- xtable(incWK2[,c(1:3,6)],
                     caption="Median Incomes by year, In the Labor Force",
                     display=c("d","d","f","d","f"))
digits(tabMedIncWK) <- 2
print(tabMedIncWK, 
      type=targetType,
      tabular.environment="longtable",
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(incWK2),by=1)),
      comment=FALSE,
      floating=FALSE
      )
```
    
In analyzing median incomes by race, the trend lines change depending on whether or not you include *all members* in the sample of the subcategory, versus using *only those in the labor force*. Here we compare Black and white workers. Proportions used in all tables are calculated using the person weights generated by the Census bureau. The black trend lines on the subsequent graphs indicate the aggregate median income, either for all individuals or all those in the labor force, depending on the relevant graph.  
\ \     
```{r, echo=FALSE}
plotMedIncBWAll
plotMedIncBWWk
```
\ \ 
```{r, echo=FALSE, results="asis"}
tabMedIncBWWk <- xtable(incBlackWhite4[,c(1:4,6,9)],
                     caption="Median Incomes for Black and White Respondents, \n In the labor force",
                     display=c("d","d","s","f","f","f","f"))
digits(tabMedIncBWWk) <- c(0,0,0,0,2,3,2)
print(tabMedIncBWWk, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(incBlackWhite4),by=2)),
      comment=FALSE
      )
```
\ \ 
Here we compare white and non-white workers^["White" workers were coded as having responded only with "White" as their race. Everyone else was coded as "Non-white", including multiracial respondents who may have included "White" as part of their race.], using the same distinction between all respondents and those included in the labor force.  

\ \ 

```{r, echo=FALSE}
plotIncWNAll
```
\ \ 
```{r, echo=FALSE}
plotIncWNWk
```
\ \ 
```{r, echo=FALSE, results="asis"}
tabMedIncWNWk <- xtable(incWhiteNon4[,c(1:3,6)],
                     caption="Median Incomes for White and Non-white Respondents, \n In the labor force",
                     display=c("d","d","s","f","f"))
digits(tabMedIncWNWk) <- 2
print(tabMedIncWNWk, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(incWhiteNon4),by=2)),
      comment=FALSE
      )
```
\ \ 

We see the same thing when we compare median incomes of all women, versus only those in the labor force.

\ \ 
``` {r, echo=FALSE}
plotIncGenderAll
```

\ \ 

``` {r, echo=FALSE}
plotIncGenderWk
```

\ \ 

```{r, echo=FALSE, results="asis"}
tabMedIncGenderWk <- xtable(incGender4[,c(1:3,6)],
                     caption="Median Incomes for Men and Women, \n In the labor force",
                     display=c("d","d","s","f","f"))
digits(tabMedIncGenderWk) <- 2
print(tabMedIncGenderWk, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(incGender4),by=2)),
      comment=FALSE
      )
```
\ \ 
\ \   

##Public versus private sector workers 
Analysis of public sector workers and private sector workers^[This was from a variable generated using the "classwkr" variable, as:  
*"Public sector"*- "Armed forces", "Federal government employee", "Local government employee", "State government employee", "Wage/salary, government".   
*"Private sector"* - "Self-employed", "Self-employed, not incorporated", "Self-employed, incorporated", "Works for wages or salary", "Wage/salary, private", "Private, for profit", "Private, nonprofit", "Unpaid family worker".], broken out by various sub-groupings.  
\ \   

First, analyzing median incomes for public versus private sector workers, analyzed all together and by sub-groupings.  
\ \ 
```{r, echo=FALSE} 
plotMedIncPubPriv
```

\ \ 

```{r, echo=FALSE} 
plotPropPubPriv
```
\ \ 
```{r, echo=FALSE, results="asis"}
tabPubPriv <- xtable(incPubPriv2[,c(1:4,6,9)],
                     caption="Share of workers in Public and Private sectors, by year",
                     display=c("d","d","s","f","f","f","f"))
digits(tabPubPriv) <- c(0,0,0,0,2,3,2)
print(tabPubPriv, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(incPubPriv2),by=2)),
      comment=FALSE
      )
```

\ \ 

For Black and white workers.  

\ \ 

```{r, echo=FALSE}
plotMedIncPubPrivBW

```

\ \ 

```{r, echo=FALSE}
plotPropPubPrivBW
```
\ \ 
```{r, echo=FALSE, results="asis"}
tabBWPubPriv <- xtable(incBWPubPriv2[,c(1:4,7,10)],
                     caption="Share of Black and White workers in Public and Private sectors, by year",
                     display=c("d","d","s","f","f","f","f"))
digits(tabBWPubPriv) <- c(0,0,0,0,2,3,2)
print(tabBWPubPriv, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(incBWPubPriv2),by=4)),
      comment=FALSE
      )
```

\ \ 

For white and non-white workers.

\ \ 

```{r, echo=FALSE}
plotMedIncPubPrivWN
```

\ \ 

```{r, echo=FALSE, results="asis"}
tabWNPubPriv <- xtable(incWNPubPriv2[,c(1:5,7,10)],
                     caption="Share of White and Non-white workers \n in Public and Private sectors, by year",
                     display=c("d","d","s","s","f","f","f","f"))
digits(tabWNPubPriv) <- c(0,0,0,0,0,2,3,2)
print(tabWNPubPriv, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(incWNPubPriv2),by=4)),
      comment=FALSE
      )
```
\ \ 
```{r, echo=FALSE}
plotMedIncPubPrivMF
```
\ \ 
```{r, echo=FALSE, results="asis"}
tabMFPubPriv <- xtable(incMFPubPriv2[,c(1:5,7,10)],
                     caption="Share of Male and Female Workers in Public and Private sectors, by year",
                     display=c("d","d","s","f","d","f","f","f"))
digits(tabMFPubPriv) <- c(0,0,0,0,0,2,3,2)
print(tabMFPubPriv, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(incMFPubPriv2),by=4)),
      comment=FALSE
      )
```
\ \ 
\ \   

##Healthcare coverage  

Measures proportion of workers covered by employer group healthcare plans.^["Covered", includes those who responded: "Yes", "Yes, paid for part", and "Yes, paid for all"] Data range from 1980-2014.

```{r, echo=FALSE}
plotHealthIns
```
\ \ 
```{r, echo=FALSE}
plotHealthInsPubPriv
```
\ \ 
```{r, echo=FALSE, results="asis"}
tabHealthInsPubPriv <- xtable(healthInsPubPriv[,c(1:4,6)],
                     caption="Share Covered by Employer's Group Healthcare plan, by Public and Private Sector",
                     display=c("d","d","s","s","d","f"))
digits(tabHealthInsPubPriv) <- 3
print(tabHealthInsPubPriv, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(healthInsPubPriv),by=4)),
      comment=FALSE
      )
```
\ \ 
```{r, echo=FALSE}
plotHealthInsWN
```
\ \ 
```{r, echo=FALSE, results="asis"}
tabHealthInsWN <- xtable(healthInsWN[,c(1:4,6)],
                     caption="Share Covered by Employer's Group Healthcare plan, \n by White and Non-white Workers",
                     display=c("d","d","s","s","d","f"))
digits(tabHealthInsWN) <- 3
print(tabHealthInsWN, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(healthInsWN),by=4)),
      comment=FALSE
      )
```    
\ \ 

##Pensions Provided by Employers  
These figures show the share of workers who receive a pension from their employer. Responses include those workers whose employers offer a pension plan, but they may not receive it.  
\ \ 
For all workers:  
\ \ 
```{r, echo=FALSE}
plotPensionCov
```

\ \ 
For public and private sector workers:  
\ \ 
```{r, echo=FALSE}
plotPensionPubPriv
```
\ \     
```{r, echo=FALSE, results="asis"}
tabPensionPubPriv <- xtable(pensionPubPriv[,c(1:4,6)],
                     caption="Receive Pension from employer, \n Public and Private Sector Workers",
                     display=c("d","d","s","s","d","f"))
digits(tabPensionPubPriv) <- 3
print(tabPensionPubPriv, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(pensionPubPriv),by=6)),
      comment=FALSE
      )
```

\ \ 

For white and non-white workers:

\ \ 

```{r, echo=FALSE}
plotPensionWN
```
\ \ 
```{r, echo=FALSE, results="asis"}
tabPensionWN <- xtable(pensionWN[,c(1:4,6)],
                     caption="Receive Pension from employer, \n White and Non-white Workers",
                     display=c("d","d","s","s","d","f"))
digits(tabPensionWN) <- 3
print(tabPensionWN, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(pensionWN),by=6)),
      comment=FALSE
      )
```
\ \ 
\ \ 

##Cleaners, healtchare workers, and sanitation workers
Analysis of these three categories of workers^[*"Cleaners"* - "Housekeepers, maids, butlers, stewards, and lodging quarters cleaners", "Private household cleaners and servants", "Private household workers allocated", "Supervisors of cleaning and building service", "Janitors", "Laundry workers", "Vehicle washers and equipment cleaners";  
*"Sanitation workers"* - "Garbage and recyclable material collectors";  
*"Healthcare Workers"* - "Nursing aides, orderlies, and attendants", "Other health and therapy", "Registered nurses", "Licensed practical nurses"], based on occupation classifications, unified to the 1990 definitions, per the census.  
\ \   

Share of white and non-white workers in the job categories^[Due to the small number of cases that fall into the *"Sanitation workers"* category, which creates noisy plots, they are not included in the graph. Raw numbers are included in the tables though.]:

\ \ 

```{r, echo=FALSE}
plotCleanersWN
```
\ \ 
```{r, echo=FALSE, results="asis"}
tabCleanersWN <- xtable(cleanersWN[,c(1:4,6)],
                     caption="Share Covered by Employer's Group Healthcare plan, by White and Non-white Workers",
                     display=c("d","d","s","s","d","f"))
digits(tabCleanersWN) <- 3
print(tabCleanersWN, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(cleanersWN),by=6)),
      comment=FALSE
      )
```  

\ \ 

Median incomes for the job categories:

\ \ 

```{r, echo=FALSE}
plotIncCleaners
```
\ \ 
```{r, echo=FALSE, results="asis"}
tabIncCleaners <- xtable(incCleaners2[,c(1:4,7)],
                     caption="Median Incomes for Jobs",
                     display=c("d","d","s","f","d","f"))
digits(tabIncCleaners) <- 3
print(tabIncCleaners, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(incCleaners2),by=3)),
      comment=FALSE
      )
```
\ \ 
Median incomes for the jobs categories in the public and private sectors:   
\ \ 
```{r, echo=FALSE}
plotIncCleanerPubPriv
```
\ \ 
```{r, echo=FALSE, results="asis"}
tabIncCleanersPubPriv <- xtable(incCleaners4[,c(1:5,8)],
                     caption="Median Incomes for Jobs in the Public and Private Sectors",
                     display=c("d","d","s","s","f","d","f"))
digits(tabIncCleanersPubPriv) <- 3
print(tabIncCleanersPubPriv, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(incCleaners4),by=6)),
      comment=FALSE
      )
```

\ \ 

Healthcare coverage for the job categories:

\ \ 

```{r, echo=FALSE}
plotHealthInsCleaners
```
\ \ 
```{r, echo=FALSE, results="asis"}
tabHealthInsCleaners <- xtable(healthInsCleaners[,c(1:4,6)],
                     caption="Share Covered by Employer's Group Healthcare plan, \n by Job Categories",
                     display=c("d","d","s","s","d","f"))
digits(tabHealthInsCleaners) <- 3
print(tabHealthInsCleaners, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(healthInsCleaners),by=6)),
      comment=FALSE
      )
```

\ \ 

Pension coverage for the job categories:

\ \ 

```{r, echo=FALSE}
plotPensionCleaners
```
\ \ 
```{r, echo=FALSE, results='asis'}
tabPensionCleaners <- xtable(pensionCleaners[,c(1:4,6)],
                     caption="Receive Pension from employer, \n by Job Categories",
                     display=c("d","d","s","s","d","f"))
digits(tabPensionCleaners) <- 3
print(tabPensionCleaners, 
      type=targetType,
      tabular.environment="longtable",
      floating=FALSE,
      caption.placement="top", 
      hline.after=c(-1,seq(0,to=nrow(pensionCleaners),by=9)),
      comment=FALSE
      )
```    
